### 2018-06-12 express2018/formal.tex

\section{Formalization}

We highlight here a formalization of CCS
in the HOL theorem
prover (HOL4) \cite{slind2008brief},
including the new concepts and theorems proposed in the first half of
this paper.
%  The main purpose is to convince the readers that, there's no flaw
%  in the informal proofs. 
The whole formalization can be found 
in \cite{Tian:2017wrba}  or in 
 in the official example folder of HOL4 source
code\footnote{\url{https://github.com/HOL-Theorem-Prover/HOL}}. The
work consists of about 20,000 lines of proof scripts in Standard ML.
%  This
%  work is based on a formalization of CCS in HOL theorem
% prover (HOL4) \cite{slind2008brief}, described in second author's MSc thesis
% \cite{Tian:2017wrba} and now lives in the official examples folder of HOL4 source
% code\footnote{\url{https://github.com/HOL-Theorem-Prover/HOL}}.

@@

\finish{below: to be reformulated (by Davide) later} 
We will not repeat all definitions and theorems of CCS again by their
formalized versions. Instead, we just focus on several highlights in
this work, i.e.
\begin{enumerate}
\item The formalization of semantics context in $\lambda$-expressions and the theory of
  (pre)congruence for CCS;
\item The definition and uses of trace in the proof of unique solution of
  contractions theorem;
\item The proof of the hard version of ``coarsest congruence
  contained in $\approx$'' theorem.
\end{enumerate}

In contrast with the theorems in the first part of the paper, in the
formalization we consider only single equations/contractions. 
This considerably simplifies the  reading of the HOL scripts. 
The extension of the work to multiple equations/contractions is
discussed in Section~\ref{s:}. \finish{to be finished}

@@

% We have only formally verified simplified forms of the unique
% solution theorems introduced in the first part. This is to not only
% save the formalization efforts, but also to make us focusing on the most
% important part of the proof. The most important simplification is
% that, we have only proved all unique solution theorems with only single-variable equations
% (contractions).

>>>>

\section{Formalisation}

We highlight here a formalisation of CCS
in the HOL theorem
prover (HOL4) \cite{slind2008brief},
including the new concepts and theorems proposed in the first half of
this paper.
%  The main purpose is to convince the readers that, there's no flaw
%  in the informal proofs. 
The whole formalisation can be found 
in \cite{Tian:2017wrba}  or in 
 in the official example folder of HOL4 source
code\footnote{\url{https://github.com/HOL-Theorem-Prover/HOL}}. The
work consists of about 20,000 lines of proof scripts in Standard ML.
%  This
%  work is based on a formalisation of CCS in HOL theorem
% prover (HOL4) \cite{slind2008brief}, described in second author's MSc thesis
% \cite{Tian:2017wrba} and now lives in the official examples folder of HOL4 source
% code\footnote{\url{https://github.com/HOL-Theorem-Prover/HOL}}.

@@

\finish{below: to be reformulated (by Davide) later} 
We will not repeat all definitions and theorems of CCS again by their
formalized versions. Instead, we just focus on several highlights in
this work, i.e.
\begin{enumerate}
\item The formalisation of semantics context in $\lambda$-expressions and the theory of
  (pre)congruence for CCS;
\item The definition and uses of trace in the proof of unique solution of
  contractions theorem;
\item The proof of the hard version of ``coarsest congruence
  contained in $\approx$'' theorem.
\end{enumerate}

In contrast with the theorems in the first part of the paper, in the
formalisation we consider only single equations/contractions. 
This considerably simplifies the  reading of the HOL scripts. 
The extension of the work to multiple equations/contractions is
discussed in Section~\ref{s:}. \finish{to be finished}

@@

% We have only formally verified simplified forms of the unique
% solution theorems introduced in the first part. This is to not only
% save the formalisation efforts, but also to make us focusing on the most
% important part of the proof. The most important simplification is
% that, we have only proved all unique solution theorems with only single-variable equations
% (contractions).

### 2018-06-12 express2018/intro.tex

The purpose of this paper is twofold. 
On the one hand, we refine the contraction proof technique by moving to the 
\emph{rooted contraction} \cite{Tian:2017wrba}, that is, the coarsest (pre)congruence contained in the contraction
preorder.  Using rooted contraction one obtains an unique-solution theorem that is valid for
\emph{rooted bisimilarity} (hence also for bisimilarity itself).  
On the other hand, we provide a  
 comprehensive formalization  of the core of the theory of CCS 
 in the HOL
theorem prover (HOL4). The formalization  includes the proofs of
Milner's 3 ``unique solution of equations'' theorems and
contractions discussed in the present paper, but is not limited to it (partly because such
theorems rely on a number of more fundamental results):
indeed the formalization encompasses the basic properties of strong and weak
bisimilarity (e.g. the fixed-point and substitutivity properties), 
their algebraic theory, various versions of ``bisimulation up to''
techniques (e.g., bisimulation up-to expansion),
the basic properties  of rooted bisimilarity. Considering the relationship between bisimilarity and rooted
bisimilarity, the formalization includes the proof that the latter is the coarsest
congruence included in the former, for which two proofs are formalized: one as in Milner's
book,  requiring the hypothesis that  no processes can use all labels; the other without
such hypothesis, essentially formalising van Glabbek's paper \cite{vanGlabbeek:2005ur}.
Similar theorems are proved for rooted contractions wrt the contraction preorder.

>>>>

The purpose of this paper is twofold. 
On the one hand, we refine the contraction proof technique by moving to the 
\emph{rooted contraction} \cite{Tian:2017wrba}, that is, the coarsest (pre)congruence contained in the contraction
preorder.  Using rooted contraction one obtains an unique-solution theorem that is valid for
\emph{rooted bisimilarity} (hence also for bisimilarity itself).  
On the other hand, we provide a  
 comprehensive formalisation  of the core of the theory of CCS 
 in the HOL
theorem prover (HOL4). The formalisation  includes the proofs of
Milner's 3 ``unique solution of equations'' theorems and
contractions discussed in the present paper, but is not limited to it (partly because such
theorems rely on a number of more fundamental results):
indeed the formalisation encompasses the basic properties of strong and weak
bisimilarity (e.g. the fixed-point and substitutivity properties), 
their algebraic theory, various versions of ``bisimulation up to''
techniques (e.g., bisimulation up-to expansion),
the basic properties  of rooted bisimilarity. Considering the relationship between bisimilarity and rooted
bisimilarity, the formalisation includes the proof that the latter is the coarsest
congruence included in the former, for which two proofs are formalized: one as in Milner's
book,  requiring the hypothesis that  no processes can use all labels; the other without
such hypothesis, essentially formalising van Glabbek's paper \cite{vanGlabbeek:2005ur}.
Similar theorems are proved for rooted contractions wrt the contraction preorder.
